An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullets

### 2021-10-07 CMPSpaper_writing/CMPSpaper_writing.Rmd
tiny edits

\wj{Additionally, we propose new metrics based on the CMPS scores and compare the new metrics with the existing metrics. 
We also introduce a principled evaluation framework of algorithmic results based on a measure called the sum of squares ratio. 
We showcase the implementation with two datasets in Hamby study (Hamby set 252 and Hamby set 44) and compare the CMPS algorithm using different sets of parameters and different metrics, following the evaluation framework based on the sum of squares ratio. }
<!-- Additionally, we apply the \pkg{cmpsR} package to two datasets in Hamby study (Hamby set 252 and Hamby set 44).  -->
The results obtained are promising:  we were both able to reproduce the results in @cmps qualitatively and achieve a clear separation between the known match (KM) comparisons and known non-match (KNM) comparisons in another bullet study.
\wj{However, the comparisons among different sets of parameters suggest that the optimal choice of the parameters varies for different datasets. 
And further research can follow the proposed evaluation framework to have a comprehensive cross-validation on parameters used the algorithm.
Furthermore, we compared our implemented CMPS algorithm with the random forest model proposed by @aoas and suggest adding the CMPS score as an additional feature in the random forest model.}

\wj{The open-source implementation of the CMPS algorithm provided by the \pkg{cmpsR} package is just one step towards the world of open science, which is important to fields like the forensic science where transparency and accuracy are critical to fair justice. 
Open-source implementations not only allow a peer-review but also facilitate further research, such as parameter cross-validations, method comparisons, and the development of statistical methods for modeling KM and KNM CMPS score distributions, which can then be used for error-rate estimations and fair applications that were called for by the PCAST [@pcast] report.
However, having open-source implementations is not enough for open science. 
In order to actually build the world of open science, many other efforts, such as open evaluation, open access publication, open educational resources, etc., are still needed. 
The database built and maintained by the National Institute of Standards and Technology is a good example of open data. 
And what we are aiming for is an open system that is able to collect open results, compare multiple algorithms using multiple datasets, and evaluate algorithmic variation and accuracy. 
The evaluation metric we proposed in this paper can be used to compare different algorithms or even different datasets and is a snippet of this open system. 
But the core of this open system is the open science culture and contributions of the community.}

>>>>

\wj{Additionally, we propose new metrics based on the CMPS scores and compare the new metrics with the existing metrics. 
We also introduce a principled evaluation framework of algorithmic results using  a measure based on the sum of squares ratio. 
We showcase the implementation with two datasets in Hamby study (Hamby set 252 and Hamby set 44) and compare the CMPS algorithm using different sets of parameters and different metrics, following the evaluation framework based on the sum of squares ratio. }
<!-- Additionally, we apply the \pkg{cmpsR} package to two datasets in Hamby study (Hamby set 252 and Hamby set 44).  -->
The results obtained are promising:  we were both able to reproduce the results in @cmps qualitatively and achieve a clear separation between the known match (KM) comparisons and known non-match (KNM) comparisons in another bullet study.
\wj{However, the comparisons among different sets of parameters suggest that the optimal choice for parameter settings vary between different datasets. We propose an evaluation framework that allows a comprehensive cross-validation of parameter settings used in the algorithm for future research. 
Comparisons of the R implementation of the CMPS algorithm with the random forest model proposed by @aoas  suggest that adding the CMPS score as an additional feature in the random forest model might add further separation between known match (KM) comparisons and known non-match (KNM) comparisons.}

\wj{The open-source implementation of the CMPS algorithm provided by the \pkg{cmpsR} package is just one step towards in the framework of open science. Open science is particularly important to fields like forensic science where transparency and accuracy are critical to fair justice. 
Open-source implementations not only allow a peer-review but also facilitate further research, such as parameter cross-validations, method comparisons, and the development of statistical methods for modeling KM and KNM CMPS score distributions, which can then be used for error-rate estimations and fair applications that were called for by the PCAST [@pcast] report.
However, having open-source implementations is not enough for open science. 
In order to actually build the world of open science, many other efforts, such as open evaluation, open access publication, open educational resources, etc., are still needed. 
The database built and maintained by the National Institute of Standards and Technology is a good example of open data. 
What we are aiming for is an open system that is able to collect open results, compare multiple algorithms using multiple datasets, and evaluate algorithmic variation and accuracy. 
The evaluation metric we proposed in this paper can be used to compare different algorithms or even different datasets and is a snippet of this open system. 
But the core of this open system is the open science culture and contributions of the community.}
